---
title: "Chicago Bulls Player Recommendation"
author: "R.Riddell"
date: "26/04/2021"
output:
  html_document:
    toc: yes
    toc_depth: 1
    toc_float:
      collapsed: no
      smooth_scroll: no
    theme: flatly
    highlight: tango
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(janitor)
library(broom)
```
# Introduction
## Background
Basketball is a team sport composed of 5 players per team and was invented around 1891 by James Naismith^(1)^. The game has evolved significantly since then and is now a very popular and lucrative sport internationally. Both players and organisations are able to earn a significant amount of money through sponsorship and media rights^(2)^. To even the financial playing field the modern salary cap was introduced in the 1984-85 season with a \$3,600,00 cap. The salary system has grown into a very complex and regulated system that ensures organisations are not able to spend any money they want on players. The NBA salary system is described as a soft cap due to a system of salary cap exceptions and is based off team’s basketball related income. Due to the NBA operating a soft cap it means there can be some range in what salary can be committed to players. Organisations needs to very strategic in the way they structure player contract in order to maximise the money they spend^(3).^ \n

## Scenario
The Chicago Bulls finished 27 from 30 teams in the most recent season, totally 22 wins from 82 games. Their point production was 104.5 points per game which ranks as the 26 out of 30 teams. As opposed to the Milwaukee Bucks who excelled with 60 wins and an average points per game of 118.1. The Chicago Bulls are coming into this season with an available \$118 million dollars for contracting players, which is 26 out of 30 compared to the rest of the league. Where the Milwaukee Bucks had \$131 million budget, showing the disparity on available budgets. \n

## Aim
The aim of this report is to flag and identify players that will enhance the probability of the Chicago Bulls winning more games than last season. \n

## Justification
The reason this report is necessary is shown by the discrepancy between what can be spent on players and what the Chicago Bulls can spend on players. This discrepancy highlights that the Chicago Bulls are not able to just recruit any player and instead must be strategic to ensure they buy players that will have a positive influence on winning for their team. \n


# Wrangling data
The descriptions and information  about the data used in this report can be found at [Data Descriptions](https://github.com/RobertRiddell/Final-Project_10157/tree/main/data-descriptions). \n

## Player Data
```{r include=F}
# Running the cleaning scripts from the funs folder to seperate the cleaninig and model analysis
source("funs/clean_player-stats.R")
obs_var1 <- dim(player_stats) 
num_players <- length(unique(player_stats$player_name))
two_team_players <- player_stats %>% 
  group_by(player_name, age) %>% 
  count() %>% 
  filter(n >1) %>% 
  pull %>%  
  length()


```
The player data consisted of `r obs_var1[2]` variables and `r obs_var1[1]` observations tabulated from the 2018/19 season. This data set included `r num_players` unique players which is assumed to be all the active players in the league.

The data relevant to players was read in and cleaned in accordance with the tidyr principles https://tidyr.tidyverse.org/articles/tidy-data.html. \n
The first issue that needed investigating was when a player had been traded during the season had both the statistics for the teams they had played for and a total row for all their statistics. There was `r two_team_players` players affected by this issue.  To resolve this issue the total row for affected players was removed from the data, leaving some players with multiple rows. \n

The next issue noted was some players had different positions listed, such as Thon Maker who was listed as a PF for Detroit and a Center for Milwaukee. It was decided that the player would be assigned the position they played the most games at. To do this the number of games was totalled for all players and a new position list was tabulated to overwrite the old positions. This function reassigned the players position on all observations to the position they played most during the season. \n

These players also had a secondary issue where they had different teams listed due to being traded. This issue was solved using the same method as the positions, where the player was assigned the team they played the most number of games from. The player data was then grouped under each name, so each player only had one observation in the dataset. \n

After grouping the data together there was some calculation errors discovered in some of the field goal percentage statistics. These issues arose as some filed goal values presented as NA and some attempt values were 0 so after the field goal calculation were completed the resultant figure was invalid. The second issue occurred through a zero-division error as there was not attempted shots so therefore the division resulted in an NA instead of a zero. During the recalculation the effective field goal statistics was also included in the data, this statistic takes into account the extra worth of a 3 point percentage. The effective field goal is calculated by using the formula  $(FG + 0.5 * 3P) / FGA$^(4)^. \n

## Team Data
```{r include=F}
# Running the cleaning scripts from the funs folder to seperate the cleaninig and model analysis
source("funs/clean_team-stats.R")
obs_var2 <- dim(team_stats)
```
The team data was sourced from three different sources, two differnet set of statistics and another seperate file that contained team payroll information. The statistical information was for each team and collected form the 2018/19 season. The two sets of team statistics were combined with the team based statistics such as pace and offensice rating removed as they were adjudged to not be applicable for player selection. The payroll data was then added to the team based data leaving the data with `r obs_var2[2]` variables and `r obs_var2[1]` observations.


# Exploratory analysis
## Distribution
```{r include=F}
tm_stats <- read_csv("data/processed/cleaned_team-stats.csv")
sum(is.na(tm_stats))

source("funs/gg_hist.R")

numeric_vars <- tm_stats %>% 
  select(where(is.numeric),- c(w,g)) %>% 
  names(.)

for (i in seq_along(numeric_vars)) {
  print(gg_hist(tm_stats, numeric_vars[i]))
}

```

After looking at a distribution of all numeric variables a selection was identified as the most normal in overall distribution. From the seven selected variables the correlation on how many games the team won versus the variable was calculated, these correlations can be seen in table 1. As seen in table 1 the statistics that returned the best correlation with winning game was points. The distribution for pts is displayed in Figure 1, the distribution is not a perfect normal distribution and could be described as bimodal with a right hand skew. Using the combination of the distribution and best correlation the point statistic has been chosen to be the best predictor of won game.
```{r echo=FALSE}
f_tr <- round(cor(tm_stats$w, tm_stats$f_tr),2)
pf <- round(cor(tm_stats$w, tm_stats$pf),2)
blk <- round(cor(tm_stats$w, tm_stats$blk),2)
trb <- round(cor(tm_stats$w, tm_stats$trb),2)
drb <- round(cor(tm_stats$w, tm_stats$drb),2)
ft_percent <- round(cor(tm_stats$w, tm_stats$ft_percent),2)
fta <- round(cor(tm_stats$w, tm_stats$fta),2)
pts <- round(cor(tm_stats$w, tm_stats$pts),2)

cor_w <- rbind(f_tr, pf, blk, trb, drb, ft_percent, fta, pts)
cor_w <- as.data.frame(cor_w)
cor_w <- rownames_to_column(cor_w, "Statistic")
cor_w <- rename(cor_w, "Correlation" = V1)
knitr::kable(cor_w,caption = "Table 1.")

dist <- ggplot(tm_stats, aes(pts)) +
  geom_histogram(colour = '#048BA8', fill = '#16DB93', alpha = 0.8, binwidth = 150) +
  labs(title = "Distribution of points",
       x = "Total Points",
       caption = "Figure 1") + 
  theme_bw()
ggsave(plot = dist, filename = "figs/figure1.jpeg")
dist
```


## Relationships
In Figure 2 the relationship between points and winning games has been presented to reinforce the decision that points will be the best indication for games won. Therefore, points will be targeted as the response variable and the other variables will be used to predict it. To standardise the variables all have been converted to per game amounts as opposed to season total amounts.
```{r echo=FALSE, message=FALSE}
scatter <- ggplot(tm_stats, aes(w, pts)) +
  geom_point() + 
  geom_smooth(method = "lm", se = F, linetype = "dashed", colour = "red") +
  labs(title = "Relatioship between Wins and Points",
       x = "Wins",
       y = "Points",
       caption = "Figure 2") +
  theme_bw()
ggsave(plot = scatter, filename = "figs/figure2.jpeg")

cor(tm_stats$w, tm_stats$pts)
scatter
```
All explanatory variables were assessed visually for their relationship with points and the correlation coefficient was calculated. Through the iterative process of modelling some variables were also removed at this stage due to multicollinearity, a poor linear relationship in the model or a slope confidence interval that straddled zero. \n

These variables were removed due to a low correlation coefficient of between -0.2 to 0.2:

* 2pa
* mp
* tov
* pf
* ft_percent
* stl
* orb

These variables were removed due to a multicollinearity:

* e_fg_percent
* ft_fga
* fg
* x3p
* x3p_ar
* ft
* fta
* drb
* fg_percent
* x2p_percent
* x3p_percent

These variables were removed due to a poor linear relationship in the model:

* blk

These variables were removed due to an unreliable confidence interval on their slope prediction:

* x3pa
* x2p
* trb
* ast
```{r include=F}
# convert stats to per game
tm_stats <- tm_stats %>% 
  mutate(mp = mp/g,
            fg = fg/g,
            fga = fga/g,
            fg_percent = fg/fga,
            x3p = x3p/g,
            x3pa = x3pa/g,
            x3p_percent = x3p/x3pa,
            x2p = x2p/g,
            x2pa = x2pa/g,
            x2p_percent = x2p/x2pa, 
            ft = ft/g,
            fta = fta/g,
            ft_percent = ft/fta,
            orb = orb/g,
            drb = drb/g,
            trb = trb/g,
            ast = ast/g,
            stl = stl/g,
            blk = blk/g,
            tov = tov/g,
            pf = pf/g,
            pts = pts/g)

team <- tm_stats$team

# variables removed as they will not add to predicting points scored
tm_stats <- tm_stats %>% 
  select(!c(g, rk,team, team_id, salary, w))

source("funs/gg_scatter.R")

numeric_vars <- tm_stats %>% 
  select(where(is.numeric),- pts) %>% 
  names(.)

for (i in seq_along(numeric_vars)) {
  print(gg_scatter(tm_stats, numeric_vars[i]))
}

tm_stats %>% 
  select(is.numeric) %>% 
  corrr::correlate() %>% 
  corrr::focus(pts) %>%
  arrange(pts) 

# remove low correlations between -0.2 and 0.2
tm_stats <- tm_stats %>% 
  select(!c(x2pa, mp, tov, pf, ft_percent, stl, orb))

# Remove obvious variables that impact each other
GGally::ggcorr(tm_stats, method = c('pairwise','spearman'), 
               size = 2 )

tm_stats <- tm_stats %>% 
  select(!c(e_fg_percent, ft_fga, fg, x3p, x3p_ar, ft, fta, drb, fg_percent, x2p_percent, x3p_percent))

# removed as it shows low linear relationship in the car::avPlots(fit)
tm_stats <- tm_stats %>% 
  select(!blk)

# Removes as the confint for the slope was -1.043498e-03	4.461750e-02 
tm_stats <- tm_stats %>% 
  select(!c(x3pa, x2p, trb, ast))

```

## Justification
As points has a reasonably normal distribution and a correlation coefficient of 0.66 it is believed that it will provide the predictor for overall games win. It is possible that another statistic that has a better distribution would model better in the data and be more predictable but as the overall goal is too increase wins it is important that the statistic chosen has strong relationship to winning 

# Data modelling and results
## Modelling
Through the previous cleaning and exploration of the data the best explanatory variables have been included in the multiple linear regression model. These variables include F-tr,ts_percent and fga. The relevant assumptions have been conducted on the data to ensure all variables meet the requirement of a multiple linear regression model.


## Assumptions
```{r include=FALSE}
# Model
fit <- lm(pts ~ ., tm_stats)
tidy(fit, conf.int = T)
summary(fit)
```

Continuous Variables

* The response variables in points are a continuous variable and all explanatory variables are also continuous.

Linear relationship

* There is an obvious linear relationship between all explanatory variables and points. This has been confirmed used the car package and avPlots function.

```{r echo=FALSE}
# Linearity
car::avPlots(fit)
```
Outliers

* The outlier plot in Figure 3 shows no obvious outliers.
```{r echo=FALSE}
#outliers
std_res <- rstandard(fit)
points <- 1:length(std_res)

outliers <- ggplot(NULL, aes(points, std_res)) +
  geom_point() +
  ylim(c(-4,4)) +
  geom_hline(yintercept = c(-3,3) , colour = "red", linetype = "dashed") + 
  labs(title = "Residual plot",
       x = "",
       y = "Residuals",
       caption = "Figure 3") +
  theme_bw()
ggsave(plot = outliers, filename = "figs/figure3.jpeg")
outliers
# no obvious outliers
```

```{r include=FALSE}
# Leverage points
tmpdf <- tm_stats 
tmpdf$hats <- hatvalues(fit)
lev_labels <- if_else(tmpdf$hats >= 0.4, paste(team), "")

ggplot(tmpdf, aes(points, hats)) +
  geom_point() +
  ggrepel::geom_text_repel(aes(label = lev_labels), colour = "navy")


# Influence
cook <- cooks.distance(fit)
cook_labels <- if_else(cook >= 1, paste(team), "")

ggplot(NULL, aes(points, cook)) +
  geom_point() +
  ggrepel::geom_text_repel(aes(label = cook_labels), colour = "navy")


```


Independence

* Using the car package and the durbinWatsonTest package a Durbin Watson statistic of 2.28 is returned which indicates a normal range and shows the model is independent.
```{r include=FALSE}
# Independence
car::durbinWatsonTest(fit)
# can range from 0 - 4
# 2 means observations are independent

```

Homoscedasticity

* To assess the homoscedasticity the predicted values are plotted against the residuals. Ideally this figure 4 will show a random spread of point which will indicate homoscedasticity.

```{r echo=FALSE}
# Homoscedasticity
res <- residuals(fit)
fitted <- predict(fit)
res_labels <- if_else(res >= .1 | res <= -.12, paste(team), "")
Homoscedasticity <- ggplot(NULL, aes(fitted, res)) +
  geom_point() +
  geom_hline(yintercept = 0, color = "red", linetype = "dashed") +
  ggrepel::geom_text_repel(aes(label = res_labels), colour = "navy") +
    labs(title = "Predicted Values against Residuals",
       x = "Preditec",
       y = "Residuals",
       caption = "Figure 4") +
  theme_bw()

ggsave(plot = Homoscedasticity, filename = "figs/figure4.jpeg")


# ideally a random spread with no clear pattern
```

Normally distributed

* To assess the distribution of the model the normality of residuals can be viewed. By looking at a histogram of the residuals and a qqplot the residuals appear normal.

```{r echo=FALSE}
# Normality of residuals

ggplot(NULL, aes(res)) +
  geom_histogram(binwidth = .07)

ggplot(NULL, aes(sample =res)) +
  stat_qq() +
  stat_qq_line()

```

## Model output
The model output shows an adjusted R sqaure value of 0.99 which indicates a very accurate model but may be an indication of overfitting in a small dataset. The slope coefficients are positive values which indicates as they increase so will points. For instance, the slope coefficient for fga is 1.23 which shows that for every 1.23 shots taken per game there will be 1 point return when all other values are held constant.
```{r include=FALSE}
tidy(fit, conf.int = T)
summary(fit)
```
```{r include=FALSE}
# multicolinearty
cor_mat <- tm_stats %>% 
  select(where(is.numeric)) %>% 
  cor(., method = 'spearman')

# select the variables that display multicolinearty above 0.8
cor_features <- caret::findCorrelation(cor_mat, cutoff = 0.8, names = T, exact = FALSE)

# Variance Inflation Factor
car::vif(fit)
sqrt(car::vif(fit))

tm_stats
```
# Player recommendation
In order to evaluate and understand players their expected point performance needs to be assessed by the constructed model. A factor that will affect a player overall expected performance is the amount of opportunity they have been given. For instance, a player that has played many more minutes should therefore have a higher number of statistics. In an attempt to mitigate this issue all players statistics can be normalised based on the average amount of minutes per game. This assumption will transform player statistics to what they would be if all players played the same minutes, obviously this assumption may not hold true based on many factors, but it is better to assess players on a more even playing field than let past opportunity be an overwhelming factor that determines future performance. 

```{r include=FALSE}
team_mins <- read_csv("data/processed/cleaned_team-stats.csv")

min_per_game <- team_mins %>% 
  group_by(team) %>% 
  summarise(ave_mins = mean(mp)/ g) %>% 
  pull %>% 
  mean

player_stats <- read_csv("data/processed/cleaned_player-stats.csv")

#players <- player_stats %>% 
  #group_by(player_name) %>% 
  #summarise(G = sum(mp)/ min_per_game,
            #f_tr = (fta/fga),
            #ts_percent = pts / ((.44*fta + fga)*2),
            #fga = sum(fga)/G,
            #x3pa = sum(x3pa)/G,
            #x2p = sum(x2p)/G,
            #trb = sum(trb)/G,
            #ast = sum(ast)/G,
            #blk = sum(blk)/G,
            #MP = sum(mp)) %>% 
  #select(-G) 

players <- player_stats %>% 
  group_by(player_name) %>%
  summarise(G = sum(mp)/ min_per_game,
            f_tr = (fta/fga),
            ts_percent = pts / ((.44*fta + fga)*2),
            fga = sum(fga)/G)


```

```{r include=FALSE}
player_stats <-  player_stats %>% 
  select(player_name, age, Pos, Tm, mp)

player_salary <- read_csv("data/raw/2018-19_nba_player-salaries.csv")

player_stats <- left_join(player_stats, player_salary, by= "player_name")

players <- mutate(players, exp_pts = predict(fit, newdata = players))

players <- players %>% 
  select(player_name, exp_pts)

player_stats <- left_join(player_stats, players, by = "player_name")

player_stats %>% 
  select(player_name, age, mp, Pos, salary, exp_pts) %>% 
  filter(mp > 300 & Pos =="PF") %>% 
  arrange(-exp_pts) 

player_stats %>% 
  select(player_name, age, mp, Pos, salary, exp_pts) %>% 
  filter(mp > 300) %>% 
  mutate(salary_points_ratio = exp_pts/salary) %>% 
  arrange(-salary_points_ratio) %>% 
  filter(Pos == "C")

player_stats %>% 
  group_by(Pos) %>% 
  filter(mp > 250) %>% 
  mutate(`Salary Rank` = rank(-salary),
         `Expected points rank` = rank(-exp_pts)) %>% 
  arrange(`Expected points rank`) %>% 
  filter(Pos == "PF")
  
# Jordan McRae	      27	333	  SG	77250	    116.136416	
# Christian Wood      23	502	  PF	1512601	  159.108397	
# Spencer Dinwiddie	  25	1914	PG	1656092	  141.702753
# Thomas Bryant	      21	1496	C	  1378242	  124.43076		
# Caris LeVert	      24	1063	SF	1702800	  126.46152	

# 77,250 + 1,512,601 + 1,656,092 + 1,378,242 + 1,702,800 = 6,326,985
# 116.136416 + 159.108397 + 141.702753 + 124.43076 + 126.46152 = 667.8398

player_stats %>% 
  select(player_name, age, mp, Pos, salary, exp_pts) %>% 
  filter(mp > 300) %>% 
  arrange(-exp_pts) 
# D'Angelo Russell	  22	2448	PG	7019698	  175.12603
# Devin Booker	      22	2242	SG	3314365	  177.381624
# Kawhi Leonard	      27	2040	SF	23114066	180.49953
# Christian Wood	     23	502	  PF	1512601	  159.10840
# Karl-Anthony Towns	23	2545	C	  7839435	  170.98054

# 7,019,698 + 3,314,365 + 23,114,066 + 1,512,601 + 7,839,435 = 42,800,165
# 175.12603 + 177.381624 + 180.49953 + 159.10840 + 170.98054 = 863.0961

player_stats %>% 
  select(player_name, age, mp, Pos, salary, exp_pts) %>% 
  filter(mp > 300) %>% 
  arrange(-exp_pts) 
# James Harden	          29	2867	PG	30570000	219.13162
# Devin Booker	          22	2242	SG	3314365	  177.381624
# Kawhi Leonard	          27	2040	SF	23114066	180.49953
# Giannis Antetokounmpo	  24	2358	PF	24157304	188.13349
# Joel Embiid	            24	2154	C	  25467250	186.03184

# 30570000 + 3314365 + 23114066 + 24157304 + 25467250 = 106,622,985
# 219.13162 + 177.381624 + 180.71463 + 188.13349 + 186.03184 = 951.3932

value_team <- player_stats %>% 
  filter(mp > 250) %>% 
  group_by(Pos) %>% 
  mutate(`Salary Rank` = rank(-salary),
         `Expected points rank` = rank(-exp_pts)) %>% 
  arrange(`Expected points rank`) %>% 
  filter(player_id == 565| 
         player_id == 150|
         player_id == 327|
         player_id == 86|
         player_id == 365) %>% 
  select(player_name, age, Pos, mp, salary, exp_pts, `Salary Rank`, `Expected points rank`) %>% 
  rename("Player Name" = player_name, "Age" = age, "Position" = Pos, "Minutes Played" = mp, "Expected Points" = exp_pts) %>% 
  arrange(Position)


highoutput_team <- player_stats %>% 
  filter(mp > 250) %>% 
  group_by(Pos) %>% 
  mutate(`Salary Rank` = rank(-salary),
         `Expected points rank` = rank(-exp_pts)) %>% 
  arrange(`Expected points rank`) %>% 
  filter(player_id == 224| 
         player_id == 67|
         player_id == 324|
         player_id == 20|
         player_id == 166) %>% 
  select(player_name, age, Pos, mp, salary, exp_pts, `Salary Rank`, `Expected points rank`) %>% 
  rename("Player Name" = player_name, "Age" = age, "Position" = Pos, "Minutes Played" = mp, "Expected Points" = exp_pts) %>% 
  arrange(Position)

moderate_team <- player_stats %>% 
  group_by(Pos) %>% 
  filter(mp > 250) %>% 
  mutate(`Salary Rank` = rank(-salary),
         `Expected points rank` = rank(-exp_pts)) %>% 
  arrange(`Expected points rank`) %>% 
  filter(player_id == 469| 
         player_id == 67|
         player_id == 324|
         player_id == 565|
         player_id == 522) %>% 
  select(player_name, age, Pos, mp, salary, exp_pts, `Salary Rank`, `Expected points rank`) %>% 
  rename("Player Name" = player_name, "Age" = age, "Position" = Pos, "Minutes Played" = mp, "Expected Points" = exp_pts) %>% 
  arrange(Position)
```
## Value for money
Using another metric derived from expected points/ salary, an understanding of the players output based on their cost can be constructed. It was possible to use this metric to select the cheapest team possible that has a moderate expected point score. This team would return an overall expected points figure of 667.89 and incur a salary of $6,326,985. This strategy certainly leaves plenty of budget for other players, but the overall expected points would not produce a significant number of wins. 
```{r echo=FALSE}
knitr::kable(value_team,caption = "Table 2: Salary and Expected Points Ranks are grouped by position")
```
## High Output
At the other end of the spectrum the best team’s expected points output would be 951.3932 but would incur a salary of $106,622,985. As players that have high output will generally record a high wage which would use up the vast majority of team’s budget. Cleary this team would not allow sufficient funds to recruit a bench unit so while the starting 5 may be excellent they are unable to play every minute of every game. With that in mind this strategy doesn’t seem a sensible or strategic decision. 
```{r echo=FALSE}
knitr::kable(highoutput_team,caption = "Table 3: Salary and Expected Points Ranks are grouped by position")
```
## Moderate team
If we approach the selection with both the overall player output and their total salary in mind, we are able to recruit a more moderate team that has an expected output of 863.10 with a salary of $42,800,165. As seen from the more efficient team Christian Wood has been selected due to his value for money but also in expected poins for his position. Kawhi Leonard is selected even though he incurs a high wage any player that is even a little bit cheaper has a signinfilat lower expected points figure, this makes him worth the high price. Karl Anthony Towns, Devin Booker and D'Angelo Russell are all selected as they are top 5 at their position for points and conversly much lower ranked for their salary. As this teams’ salary is also moderate it will allow for plenty of budget for the other 7-10 players required. There is enough salary remaining the team could afford to invest heavily in 2-3 players to make up the core of the bench unit and be good cover for injury for the starting players. These are the five player who appear as the best value for money while also being able to deliver an increased number of wins.  
```{r echo=FALSE}
knitr::kable(moderate_team,caption = "Table 4: Salary and Expected Points Ranks are grouped by position")

```

# Summary
The model constructed in this report is one that represent as very accurate, there is a concern that due to the small number of samples the model is overfitting the data and predicting the exact points which will lead to poor prediction with unfamiliar data. \n
Other limitations of using points to predict game wins is the overall correlation of .66 is reasonable but not perfect, as seen in Figure 2 there are plenty of teams that scored plenty of points and underperformed in relation to wins and vice versa some teams had lower total points but overperformed on wins. This may mean the recruited team can score a high number of total points but be unable to convert those points to won games. \n
The model also does not include any statistics outside of scoring points, as the prediction was aimed at points it logically means that the best variable will be directly related to scoring points. This will mean the model will give the best expected point rating to players that are good shooters and take lots of shots. Scoring points is an important element of basketball but not the whole picture, so it would be advisable with the remaining team budget be focused on players that ore skill lay outside of scoring points. In this way it would give the team balance and through the flow of the team’s rotations team would have both the capability to score lots of points and hold opposition teams to low scores

# Reference List
Provide a reference list of any sources you used in the development of your report and justification of your arguments. Please use the Vancouver reference style (Links to an external site.) for the reference list and in-text references. 

1.	Donald LW, Logan RG, Mokray WG. Basketball. In: Encyclopedia Britannica. 2021.
2.	Reiff N. How the NBA makes money [Internet]. Investopedia.com. 2020 [cited 2021 May 3]. Available from: https://www.investopedia.com/articles/personal-finance/071415/how-nba-makes-money.asp
3.	CBA Breakdown [Internet]. Cbabreakdown.com. [cited 2021 May 3]. Available from: https://cbabreakdown.com/salary-cap-overview
4.  Page B, Page B. Glossary | Basketball-Reference.com [Internet]. Basketball-Reference.com. 2021 [cited 3 May 2021]. Available from: https://www.basketball-reference.com/about/glossary.html#:~:text=eFG%25%20%2D%20Effective%20Field%20Goal%20Percentage,a%202%2Dpoint%20field%20goal

 